Stage-Splitting a Modal Language

### 2014-10-17 paper/umut.tex

...To be added to the paper...

@@

\newcommand{\drun}[2]{\lVert{#2}\rVert_{#1}}
\paragraph{Partial Evaluation.}

The following definitions are adapted from the partial evaluation
book.  For simplicity and uniformity, we assume that data and programs
are draws from the same set and that ill-behaved programs and
non-terminating programs all return $bottom$.

@@

Ershov also introduced a richer form of transformation, {\em mixed
  computation} that generalizes partial evaluation.  While in some
literature, mixed computation is characterized as similar to partial
evaluation, it is significantly more general, at least by definition.

@@

Partial evaluation is trivially a special case of mixed computation,
because it allows a transformation of the program only on the static
data, and because it allows only a simple transformation on data (the
extraction of the dynamic data).

>>>>

...To be added to the paper...
(TODO: make a pass to make consistent the terminology. it is a bit complex.)

@@

\newcommand{\drun}[2]{\lVert{#2}\rVert_{#1}}

\subsection{Mixed Computation, Partial Evaluation, and Data Specialization.}

The following definitions are adapted from the partial evaluation
book, the cited papers, and Malmkjaer thesis.  For simplicity and
uniformity, we assume that data and all programs are drawn from the
same set and that ill-behaved programs and non-terminating programs
all return $bottom$.

@@

Data specialization (Barzdin and Bulyonkov 88) is a generalization of
partial evaluation.

Definition [Data Specialization].  A data specializer for a language
$P$ consist of two algorithms $ds_p, ds_d \in P$ for specializing the
data and the program such that for any program $p \in P$ and any
static and dynamic data $d_s, d_d \in P$, the code-specialized program
when run with static-data-specialized program behaves the same as the
original program run with the original data, i.e.,

\[
\drun{P}{p} (d_s, d_d) = \drun{P}{\drun{P}{ds_p} (p)} (\drun{P}{ds_d}
(p,d_s), d_d)).
\]

Seeing that partial evaluation is a special case of data
specialization is not trivial, as it requires using an interpreter for
the language $P$ (written in $P$).  Specifically, we can use a partial
evaluator in place of data specializer $ds_d$ and an  interpreter for $P$ in
place of the program specializer $ds_p$.

Both partial evaluation and data specialization are special cases of
mixed computation, which allows both the program and the data to be
specialized with respect to each other.  Ershov introduced {\em mixed
  computation}.

@@

Partial evaluation is trivially a special case of mixed computation,
because it allows a transformation of the program only on the static
data, and because it allows only a simple transformation on data (the
extraction of the dynamic data).  Similarly data specialization is a
special case of mixed computation because it allows the program to be
specialized in the absence of data and only the static portion of the
data to be specialized with respect to the program.

\subsection{Techniques}

Polyvariant.

Call Unrolling.

Elimination of duplicate work.
